---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "90%",
  fig.align = "center",
  dev = "ragg_png", # <3
  dpi = 132
)

library(ggpointless)
```

# ggpointless

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/flrd/ggpointless/branch/main/graph/badge.svg)](https://app.codecov.io/gh/flrd/ggpointless?branch=main)
<!-- badges: end -->

`ggpointless` adds visual sugar to your plots. The package provides a simple point layer to emphazise some observations.

## Installation

You can install the development version of ggpointless from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("flrd/ggpointless")
```

## Hello world

The one `geom` in this package is `geom_pointless`. It behaves like `geom_point()` but, by default, `geom_pointless()` adds only a single point at the last observation. 

```{r default}
library(ggpointless)
# set theme
theme_set(theme_minimal() +
            theme(legend.position = "bottom"))

# dummy data
x <- seq(-2 * pi, 2 * pi, length.out = 500)
y <- sin(1/4 * x) + sin(1/2 * x) + sin(x)
df1 <- data.frame(
  var1 = x,
  var2 = y
)

# ggplot object
p <- ggplot(df1, aes(x = var1, y = var2))
p +
  geom_pointless()
```

You can set the argument `location` to `"first"`, `"last"`, `"minimum"`, `"maximum"`, and `"all"`, where `"all"` is just shorthand to select `"first"`, `"last"`, `"minimum"` and `"maximum"`.  

```{r all}
p +
  geom_pointless(location = "all")
```

As you can see, `geom_pointless()` is not terribly useful on its own (and here I am thinking about a better package name) but it is when it teams up with `geom_line()`, hopefully. `geom_pointless()` understands the same arguments as `geom_point()` and it computes one additional variable -- `location` -- that you can map to an aesthetic, e.g. `colour`.  

```{r line}
p +
  geom_line() +
  geom_pointless(
    aes(colour = after_stat(location)),
    location = "all",
    size = 3
    )
```

`geom_pointless()` treats the observations in the order in which they appear in your data (like `geom_path()` does.) Example data taken from the [`geomtextpath`](https://github.com/AllanCameron/geomtextpath/#using-geomtextpath) package:

```{r spiral}
x <- seq(5, -1, length.out = 1000) * pi
spiral <- data.frame(var1 = sin(x) * 1:1000, 
                     var2 = cos(x) * 1:1000)

ggplot(spiral, aes(var1, var2)) +
  geom_path() +
  geom_pointless(
    aes(colour = after_stat(location)),
    location = "all") +
  coord_equal()
```

As you see from the last example, `"first"` and `"minimum"` overlap, and `"first"` wins over `"minimum"`. The order in which points are plotted (if specified together) from top to bottom is `"first"`, `"last"`, `"minimum"`, then `"maximum"`. 


### Facets
Since we make use the magic of [`ggplot2`](https://ggplot2.tidyverse.org/), the geom can of course be used for faceting. The following example is taken from [Stack Overflow](https://stackoverflow.com/q/29375169/8583393)

```{r stackoverflow}
# https://stackoverflow.com/q/29375169/8583393

ggplot(economics_long,
       aes(x = date, y = value)) + 
  geom_line() +
  facet_wrap(~variable, ncol = 1, scales = 'free_y') +
  geom_pointless(
    aes(colour = after_stat(location)),
    location = c("minimum", "maximum")
    )
```


## Why ggpointless?

Fair question. 

## Related work

The package [`ggpmisc`](https://exts.ggplot2.tidyverse.org/ggpmisc.html) provides the stat `stat_peaks`, which calls the `peaks()` function from the [`splus2r`](https://github.com/spkaluzny/splus2r) package. With these functions / stats you are able to find and hence highlight _local_ minima and maxima too, besides global extrema.

