---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-", 
  out.width = "100%",
  fig.align = "center",
  fig.height = 3,
  dev = "ragg_png", # <3
  dpi = 132
)

library(ggpointless)
```

# ggpointless

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/flrd/ggpointless/branch/main/graph/badge.svg)](https://app.codecov.io/gh/flrd/ggpointless?branch=main)
[![R-CMD-check](https://github.com/flrd/ggpointless/workflows/R-CMD-check/badge.svg)](https://github.com/flrd/ggpointless/actions)
<!-- badges: end -->

The goal is to emphasize certain observations with an accent to provide some additional context.

## Installation

You can install the development version of `ggpointless` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("flrd/ggpointless")
```

## Usage

There are two functions in the `ggpointless` package: `geom_pointless()`, which is powered by `stat_pointless()`. Both functions add a point layer to a `ggplot` object by default. In addition to `geom_point()` the function have an additional `location` argument. You can set it to `"first"`, `"last"` (default), `"minimum"`, `"maximum"`, and `"all"`, where `"all"` is just shorthand to select `"first"`, `"last"`, `"minimum"` and `"maximum"`.  

See the `vignette("ggpointless")` for more details.


```{r hello_world}
x <- seq(-pi, pi, length.out = 100)
y <- outer(x, 1:5, function(x, y) sin(x*y))

df1 <- data.frame(
  var1 = x,
  var2 = rowSums(y)
)

ggplot(df1, aes(x = var1, y = var2)) +
  geom_line() +
  geom_pointless(aes(colour = after_stat(location)),
                 location = "all",
                 size = 3) +
  scale_color_manual(values = c('#f4ae1b', '#d77e7b', '#a84dbd', '#311dfc')) +
  theme_minimal()
```


## Data

The `ggpointless` package contains two data sets:

1. `co2_ml` : [CO~2~ records taken at Mauna Loa](https://gml.noaa.gov/ccgg/trends/data.html)
2. `covid_vac` : [COVID-19 Cases and Deaths by Vaccination Status](https://covid.cdc.gov/covid-data-tracker/#rates-by-vaccine-status)

See the `vignette("examples")` for possible use cases.

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/flrd/ggpointless/blob/master/CONDUCT.md). By participating in this project you agree to abide by its terms.
