---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  # fig.width = 3,
  fig.height = 3, 
  out.width = "90%",
  fig.align = "center",
  dev = "ragg_png", # <3
  dpi = 132
)

library(ggpointless)
```

# ggpointless

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/flrd/ggpointless/branch/main/graph/badge.svg)](https://app.codecov.io/gh/flrd/ggpointless?branch=main)
[![R-CMD-check](https://github.com/flrd/ggpointless/workflows/R-CMD-check/badge.svg)](https://github.com/flrd/ggpointless/actions)
<!-- badges: end -->

The package provides a simple point layer to emphazise some observations in your data.

## Installation

You can install the development version of ggpointless from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("flrd/ggpointless")
```

## Usage

There are two functions in the `ggpointless` package: `geom_pointless()`, which is powered by `stat_pointless()`. Both functions add a layer to a `ggplot` object; a point layer by default. 


```{r hello_world}
# dummy data
x <- seq(-pi, pi, length.out = 100)
y <- outer(x, 1:5, FUN = \(x, y) sin(x*y)) |> 
  rowSums()

df1 <- data.frame(
  var1 = x,
  var2 = y
)

# plot
ggplot(df1, aes(x = var1, y = var2)) +
  geom_line() +
  geom_pointless(
    aes(colour = after_stat(location)),
    location = "all",
    size = 2.5) +
  theme_void()
```


## Motivation

Laziness. I found myself trying to recreate this fantastic plot from Gregor Aisch on [Carbon dioxide concentration over time](https://blog.datawrapper.de/weekly-chart-carbon-dioxide/). 


More examples and details can be found in the package's [vignette]().

## Related work

The package [`ggpmisc`](https://exts.ggplot2.tidyverse.org/ggpmisc.html) provides the stat `stat_peaks`, which calls the `peaks()` function from the [`splus2r`](https://github.com/spkaluzny/splus2r) package. With these functions / stats you are able to find and hence highlight _local_ minima and maxima too, besides global extrema.

