---
title: "Introduction to ggpointless"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ggpointless}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  dev       = "ragg_png",
  fig.align = "center",
  out.width = "100%"
)
library(ggplot2)
library(ggpointless)

cols <- c("#f4ae1b", "#d77e7b", "#a84dbd", "#311dfc")
theme_set(theme_minimal())
```

`ggpointless` is a small extension of the [`ggplot2`](https://ggplot2.tidyverse.org/)
package that provides additional layers:

1. `geom_area_fade()`
2. `geom_arch()` & `stat_arch()`
3. `geom_catenary()` & `stat_catenary()`
4. `geom_chaikin()` & `stat_chaikin()`
5. `geom_fourier()` & `stat_fourier()`
6. `geom_lexis()` & `stat_lexis()`
7. `geom_point_glow()`
8. `geom_pointless()` & `stat_pointless()`


## geom_area_fade

`geom_area_fade()` behaves like `geom_area()` but fills each group with a
linear gradient that fades from the fill colour to transparent at the baseline,
drawing the eye to the signal rather than the area underneath.

```{r area-fade-economics, fig.height = 3}
ggplot(economics, aes(date, unemploy / 1000)) +
  geom_area_fade(fill = "#311dfc") +
  labs(x = NULL, y = "Unemployed (millions)", title = "US unemployment 1967–2015")
```

Use `alpha` to set the starting opacity and `alpha_fade_to` to control where
the gradient ends. The direction can be reversed — transparent at the top,
opaque at the baseline — by swapping their values.

```{r area-fade-alpha, fig.height = 3}
ggplot(economics, aes(date, unemploy / 1000)) +
  geom_area_fade(fill = "#d77e7b", alpha = 0.9, alpha_fade_to = 0.05) +
  labs(x = NULL, y = "Unemployed (millions)")
```

Multiple groups each get their own gradient:

```{r area-fade-groups, fig.height = 3}
df <- data.frame(
  g = c("a", "a", "a", "b", "b", "b"),
  x = c(1, 3, 5, 2, 4, 6),
  y = c(2, 5, 1, 3, 6, 7)
)
ggplot(df, aes(x, y, fill = g)) +
  geom_area_fade(alpha = 0.85) +
  scale_fill_manual(values = c(a = "#f4ae1b", b = "#311dfc"))
```


## geom_arch & stat_arch

### Catenary curves and arches

A **catenary** (from the Latin *catena*, chain) is the curve formed by a
flexible chain or cable hanging freely between two fixed supports. Contrary to
a widely held belief, it is *not* a parabola — Galileo thought it was, but
Leibniz, Huygens, and Johann Bernoulli independently derived the correct
equation in 1691:

$$y = a \cosh\!\left(\frac{x}{a}\right)$$

where $a$ controls how steeply the chain curves. What makes the catenary
structurally remarkable is its **inverted form**: an inverted catenary carries
its own weight in pure compression with no bending moments — the most efficient
load path for masonry and unreinforced concrete. Antoni Gaudí famously designed
the arches of the [Sagrada Família](https://en.wikipedia.org/wiki/Sagrada_Família)
by hanging chains, photographing the inverted result, and scaling up. The
[Gateway Arch](https://en.wikipedia.org/wiki/Gateway_Arch) in St Louis (1965)
is a weighted catenary 192 metres tall.

`geom_arch()` draws these inverted catenary curves between successive points.
The shape is controlled by `arch_length` (total arc length) or `arch_height`
(vertical rise above the highest endpoint of each segment). By default the arc
length is twice the Euclidean distance.

```{r arch-basic, fig.height = 3}
df_arch <- data.frame(x = 1:5, y = c(0, 0.3, 0, 0.3, 0))

ggplot(df_arch, aes(x, y)) +
  geom_arch(arch_height = 0.6, colour = "#a84dbd") +
  geom_point(size = 3, colour = "#a84dbd")
```

### The Rice House

A real-world application is the **[Rice House](https://en.wikipedia.org/wiki/Rice_House,_Eltham)**
in Eltham (1966), whose distinctive roofline consists of shallow catenary arched
spans rising above vertical columns. The sketch below reconstructs its simplified
cross-section:

```{r rice-house, fig.height = 3}
rice_house <- data.frame(
  x = c(0,   1.5, 2.5, 3.5, 5),
  y = c(0,   1,   1,   1,   0)
)

ggplot(rice_house, aes(x, y)) +
  geom_arch(arch_height = 0.15, linewidth = 2, colour = "#5c4033") +
  geom_segment(aes(xend = x, yend = 0), linewidth = 2, colour = "#5c4033") +
  geom_hline(yintercept = 0, colour = "#4a7c59", linewidth = 5) +
  coord_equal() +
  theme_void() +
  labs(caption = "Rice House, Eltham (1966) — simplified catenary cross-section")
```

### Controlling arch height with stat_arch

`stat_arch()` exposes the underlying computation. Here it shows the effect of
different `arch_height` values between the same two endpoints:

```{r stat-arch-compare, fig.height = 3.5}
ggplot(data.frame(x = c(0, 5), y = c(0, 0)), aes(x, y)) +
  stat_arch(arch_height = 0.5, colour = "#f4ae1b") +
  stat_arch(arch_height = 1.5, colour = "#d77e7b") +
  stat_arch(arch_height = 3.0, colour = "#311dfc") +
  geom_point(size = 3) +
  labs(title = "arch_height = 0.5 (yellow), 1.5 (pink), 3.0 (blue)")
```


## geom_catenary & stat_catenary

`geom_catenary()` draws the *non*-inverted, hanging-chain form between
successive points.

```{r catenary-basic, fig.height = 3}
set.seed(1)
ggplot(data.frame(x = 1:6, y = sample(6)), aes(x, y)) +
  geom_catenary(colour = "#f4ae1b") +
  geom_point(size = 3)
```

### chain_length

By default the chain length is twice the Euclidean distance per segment. Pass
`chain_length` to set an explicit arc length — longer values make the chain sag
more. If the value is shorter than the straight-line distance a warning is
issued and a straight line is drawn instead.

```{r catenary-chain-length, fig.height = 3}
ggplot(data.frame(x = c(0, 1), y = c(1, 1)), aes(x, y)) +
  lapply(seq(1.1, 2.5, by = 0.3), function(cl) {
    geom_catenary(chain_length = cl, colour = "#311dfc", alpha = 0.55)
  }) +
  geom_point(size = 3) +
  ylim(NA, 1.05) +
  labs(title = "Increasing chain_length adds more sag")
```

### sag

`sag` sets the vertical drop below the lowest endpoint of each segment, giving
you direct control over how much the chain droops:

```{r catenary-sag, fig.height = 3}
df_sag <- data.frame(x = c(0, 2, 4, 6), y = c(1, 1, 1, 1))

ggplot(df_sag, aes(x, y)) +
  geom_catenary(sag = c(0.1, 0.5, 1.2), colour = "#d77e7b") +
  geom_point(size = 3) +
  labs(title = "sag = 0.1, 0.5, 1.2 (left to right)")
```


## geom_chaikin & stat_chaikin

Chaikin's corner-cutting algorithm smooths a polygonal path by iteratively
replacing each corner with two new points placed at a given ratio along the
adjacent edges. With `ratio = 0.25` (the default) and `iterations = 5` the
path converges to a smooth B-spline approximation.

```{r chaikin-basic, fig.height = 3}
set.seed(42)
dat <- data.frame(x = seq.int(10), y = sample(15:30, 10))

ggplot(dat, aes(x, y)) +
  geom_line(linetype = "dashed", colour = "grey60") +
  geom_chaikin(colour = "#311dfc") +
  labs(title = "Original path (dashed) vs Chaikin-smoothed (blue)")
```

### Effect of ratio

`ratio` controls how aggressively corners are cut. At `ratio = 0.5` the new
points coincide with the edge midpoints, producing the maximum rounding
possible in a single pass.

```{r chaikin-ratio, fig.height = 3}
triangle <- data.frame(x = c(0, 0.5, 1), y = c(0, 1, 0))

ggplot(triangle, aes(x, y)) +
  geom_polygon(fill = NA, colour = "grey70", linetype = "dashed") +
  geom_chaikin(ratio = 0.10, mode = "closed", colour = "#f4ae1b") +
  geom_chaikin(ratio = 0.25, mode = "closed", colour = "#d77e7b") +
  geom_chaikin(ratio = 0.50, mode = "closed", colour = "#311dfc") +
  coord_equal() +
  labs(subtitle = "ratio = 0.10 (yellow), 0.25 (pink), 0.50 (blue)")
```

### Smoothing a closed polygon with stat_chaikin

`stat_chaikin()` can be combined with any geom. Pass `geom = "polygon"` to
smooth the boundary of a filled shape:

```{r chaikin-polygon, fig.height = 3.5}
# An irregular hexagon
hex <- data.frame(
  x = c(0.0, 0.4, 1.0, 1.2, 0.9, 0.1),
  y = c(0.2, 1.0, 0.9, 0.3, -0.2, 0.0)
)

ggplot(hex, aes(x, y)) +
  geom_polygon(fill = "grey85", colour = "grey50", linetype = "dashed") +
  stat_chaikin(geom = "polygon", mode = "closed",
               fill = "#a84dbd", alpha = 0.45, colour = "#a84dbd") +
  coord_equal() +
  labs(title = "Original polygon (dashed) with smoothed fill (purple)")
```


## geom_fourier & stat_fourier

`geom_fourier()` fits a truncated **discrete Fourier transform** to the
supplied x/y data and renders the reconstructed curve. Fewer harmonics produce
a smoother, low-frequency summary; retaining all harmonics reproduces the
original signal exactly (up to interpolation artefacts).

```{r fourier-harmonics, fig.height = 3.5}
set.seed(42)
n <- 150
df_f <- data.frame(
  x = seq(0, 2 * pi, length.out = n),
  y = sin(seq(0, 2 * pi, length.out = n)) +
      0.4 * sin(3 * seq(0, 2 * pi, length.out = n)) +
      rnorm(n, sd = 0.25)
)

ggplot(df_f, aes(x, y)) +
  geom_point(alpha = 0.25, size = 0.8) +
  geom_fourier(n_harmonics = 1,  colour = "#f4ae1b") +
  geom_fourier(n_harmonics = 3,  colour = "#d77e7b") +
  geom_fourier(n_harmonics = 10, colour = "#311dfc") +
  labs(
    title = "1 harmonic (yellow), 3 (pink), 10 (blue)",
    x = NULL, y = NULL
  )
```

### Detrending

A linear drift in the data will dominate the low-frequency coefficients,
preventing the Fourier series from capturing the periodic structure. Set
`detrend = "lm"` (or `"loess"`) to subtract the trend before the transform;
it is added back afterwards so the output remains on the original scale.

```{r fourier-detrend, fig.height = 3.5}
set.seed(3)
x_d <- seq(0, 4 * pi, length.out = 100)
df_d <- data.frame(
  x = x_d,
  y = sin(x_d) + x_d * 0.4 + rnorm(100, sd = 0.2)
)

ggplot(df_d, aes(x, y)) +
  geom_point(alpha = 0.35, size = 0.8) +
  geom_fourier(aes(colour = "as is"), n_harmonics = 3, 
               linetype = "dashed") +
  geom_fourier(aes(colour = "detrend = 'lm'"), n_harmonics = 3, detrend = "lm") +
  scale_colour_manual(values = c("as is" = "#a84dbd", "detrend = 'lm'" = "#311dfc")) +
  labs(
    title   = "With and without detrending",
    x = NULL, y = NULL
  )
```

### Pitfall: irregular spacing with calendar dates

The Fourier transform (via `stats::fft()`) assumes that observations are
**evenly spaced** in time. When you hand it calendar dates directly, this
assumption is silently violated: months have 28–31 days, so consecutive
observations are not the same distance apart.

The example below maps `date` — a `Date` vector — to the x aesthetic and
asks `geom_fourier()` to reconstruct the signal with all harmonics
(`n_harmonics = NULL`). You might expect the Fourier curve to pass through
every data point, but it does not:

```{r fourier-dates-fail, fig.height = 3.5}
ggplot(head(economics, 20), aes(date, unemploy)) +
  geom_line(linetype = "dotted") +
  geom_point(size = 1) +
  geom_fourier(n_harmonics = NULL, colour = "#311dfc") +
  labs(
    title = "geom_fourier() with Date x-axis — curve misses observations",
    x = NULL, y = "Unemployed (thousands)"
  )
```

The Fourier layer converts dates to their underlying integer representation
(days since 1970-01-01), constructs a **uniform** grid of that many integer
steps, computes the inverse FFT on that grid, and draws a line through those
interpolated points. Because the grid points fall between observation dates,
the curve misses the actual data points even when every harmonic is included.

To fix this you can replace the date x-axis with a plain integer
index so that spacing is guaranteed to be uniform. You can restore
human-readable year labels with `scale_x_continuous()`:

```{r fourier-dates-fix, fig.height = 3.5}
eco <- head(economics, 20)
eco$tmp <- seq_len(nrow(eco))

# integer positions for year-start labels
yr_breaks <- which(format(eco$date, "%m") == "01")

ggplot(eco, aes(tmp, unemploy)) +
  geom_line(linetype = "dotted") +
  geom_point(size = 1) +
  geom_fourier(n_harmonics = NULL, colour = "#311dfc") +
  scale_x_continuous(
    breaks = yr_breaks,
    labels = format(eco$date[yr_breaks], "%Y")
  ) +
  labs(
    title = "Integer index x-axis — curve passes through every observation",
    x = NULL,
    y = "Unemployed (thousands)"
  )
```

The Fourier curve now interpolates the data exactly. When working with any
irregularly-spaced time series — monthly, weekly, or business-day — convert
the time variable to a sequential integer index before passing it to
`geom_fourier()` or `stat_fourier()`.


## geom_lexis & stat_lexis

`geom_lexis()` draws a 45° lifeline for each observation from its start to its
end. The required aesthetics are `x` and `xend`; `y` and `yend` are calculated
by `stat_lexis()` and represent the cumulative duration.

```{r lexis-basic, fig.height = 3.5}
df_l <- data.frame(
  key  = c("A", "B", "B", "C", "D"),
  x    = c(0, 1, 6, 5, 6),
  xend = c(5, 4, 10, 8, 10)
)

p <- ggplot(df_l, aes(x = x, xend = xend, colour = key)) +
  coord_equal() +
  scale_colour_manual(values = cols)

p + geom_lexis()
```

When there is a gap between two events of the same cohort, a horizontal dotted
segment bridges the gap. Set `gap_filler = FALSE` to hide it.

```{r lexis-gap, fig.height = 3.5}
p + geom_lexis(gap_filler = FALSE)
```

### Using after_stat(type) and custom point styling

The computed variable `type` takes the value `"solid"` for 45° lifelines and
`"dotted"` for horizontal gap-fillers. Map it to `linetype` to make the
distinction explicit, and use `point_colour` to style the endpoint dot
independently of the lifeline colour.

```{r lexis-type, fig.height = 3.5}
p +
  geom_lexis(
    aes(linetype = after_stat(type)),
    point_colour = "black",
    shape        = 21,
    fill         = "white",
    size         = 2.5,
    stroke       = 0.8
  ) +
  scale_linetype_identity()
```

### Date and POSIXct classes

`geom_lexis()` works with Date and POSIXct objects as well as numerics. The
y-axis shows duration in the native unit of the scale (days for Date, seconds
for POSIXct).

```{r lexis-dates, fig.height = 3.5}
df_dates <- data.frame(
  key   = c("A", "B"),
  start = c(2019, 2021),
  end   = c(2022, 2022)
)
df_dates[, c("start", "end")] <- lapply(
  df_dates[, c("start", "end")],
  function(i) as.Date(paste0(i, "-01-01"))
)

ggplot(df_dates, aes(x = start, xend = end, group = key)) +
  geom_lexis() +
  scale_y_continuous(
    breaks = 0:3 * 365.25,
    labels = function(i) paste0(floor(i / 365.25), " yr")
  ) +
  coord_fixed() +
  labs(y = "Duration")
```


## geom_point_glow

`geom_point_glow()` is a drop-in replacement for `geom_point()` that adds a
radial gradient glow behind each point using `grid::radialGradient()`. It
shines brightest on dark backgrounds.

### Ursa Major

The constellation Ursa Major, drawn with star positions in right ascension and
declination:

```{r ursa-major, fig.height = 5, fig.width = 7.5}
stars_base <- data.frame(
  star  = c(
    "Dubhe", "Merak", "Phecda", "Megrez", "Alioth", "Mizar", "Alkaid",
    "Muscida", "Talitha", "Tania Borealis", "Tania Australis",
    "Alula Borealis", "Alula Australis", "Upsilon", "Phi", "Theta", "Chi", "Psi"
  ),
  ra_h  = c(
    11.0, 11.0, 11.9, 12.3, 12.9, 13.4, 13.8,
     8.5,  9.0, 10.3, 10.3, 11.3, 11.3,
     9.8,  9.9,  9.5, 11.8, 11.2
  ),
  dec_d = c(
    62, 56, 54, 57, 56, 55, 49,
    61, 48, 43, 41, 33, 31,
    59, 54, 52, 48, 45
  )
)

path1 <- c("Alkaid", "Mizar", "Alioth", "Megrez", "Phecda", "Merak", "Dubhe", "Megrez")
path2 <- c("Dubhe", "Muscida", "Upsilon", "Phi", "Theta", "Talitha")
path3 <- c("Phecda", "Chi", "Psi", "Tania Australis", "Tania Borealis")
path4 <- c("Chi", "Alula Borealis", "Alula Australis")

make_path <- function(star_names, group_id) {
  df        <- stars_base[match(star_names, stars_base$star), ]
  df$group  <- group_id
  df
}

plot_data <- rbind(
  make_path(path1, 1),
  make_path(path2, 2),
  make_path(path3, 3),
  make_path(path4, 4)
)
plot_data$x <- -plot_data$ra_h
plot_data$y <-  plot_data$dec_d

ggplot(plot_data, aes(x = x, y = y)) +
  geom_path(aes(group = group), colour = "#f4ae1b", linewidth = 0.6) +
  geom_point_glow(
    data        = stars_base,
    aes(x = -ra_h, y = dec_d),
    colour      = ggplot2::alpha("#f4ae1b", alpha = .5),
    size        = 2,
    glow_colour = "#f4ae1b",
    glow_alpha  = 0.75
  ) +
  geom_text(
    data  = stars_base,
    aes(x = -ra_h, y = dec_d, label = star),
    colour        = "#bbccdd",
    vjust         = -1.2,
    size          = 2.5,
    check_overlap = TRUE
  ) +
  scale_x_continuous(
    breaks = seq(-14, -8, by = 1),
    labels = \(x) paste0(abs(x), "h")
  ) +
  scale_y_continuous(labels = \(x) paste0(x, "°")) +
  labs(
    title = "Ursa Major",
    x     = "Right Ascension",
    y     = "Declination"
  ) +
  theme(
    panel.background = element_rect(fill = "#0d1b2a", colour = NA),
    plot.background  = element_rect(fill = "#0d1b2a", colour = NA),
    panel.grid.major = element_line(colour = "#1e2f3e", linetype = "dotted"),
    panel.grid.minor = element_blank(),
    text             = element_text(colour = "white"),
    axis.text        = element_text(colour = "#8899aa"),
    plot.title       = element_text(size = 14, face = "bold")
  )
```


## geom_pointless & stat_pointless

`geom_pointless()` highlights selected observations — first, last, minimum,
maximum, or all four — with points. Its name reflects that it is not
particularly useful on its own, but in conjunction with `geom_line()` and
friends, it adds useful context at a glance.

```{r pointless-basic, fig.height = 3}
x   <- seq(-pi, pi, length.out = 100)
df1 <- data.frame(var1 = x, var2 = rowSums(outer(x, 1:5, function(x, y) sin(x * y))))

p <- ggplot(df1, aes(x = var1, y = var2)) + geom_line()

p + geom_pointless(location = "all", size = 3)
```

Map the computed variable `location` to `colour` to distinguish the four roles:

```{r pointless-colour, fig.height = 3}
loc_cols <- c(
  "first"   = "#f4ae1b",
  "last"    = "#d77e7b",
  "minimum" = "#a84dbd",
  "maximum" = "#311dfc"
)

p +
  geom_pointless(
    aes(colour = after_stat(location)),
    location = "all",
    size     = 3
  ) +
  scale_colour_manual(values = loc_cols) +
  theme(legend.position = "bottom")
```

### Order and orientation

Locations are determined in data order, as `geom_path()` would — not sorted by
x like `geom_line()`. This matters when the path crosses itself:

```{r pointless-spiral, fig.height = 3.5, fig.show = 'hold'}
x       <- seq(5, -1, length.out = 1000) * pi
spiral  <- data.frame(var1 = sin(x) * 1:1000, var2 = cos(x) * 1:1000)

p_spi <- ggplot(spiral) +
  geom_path() +
  coord_equal(xlim = c(-1000, 1000), ylim = c(-1000, 1000)) +
  theme(legend.position = "none")

p_spi +
  aes(x = var1, y = var2) +
  geom_pointless(aes(colour = after_stat(location)), location = "all", size = 3) +
  labs(subtitle = "orientation = 'x'")

p_spi +
  aes(y = var1, x = var2) +
  geom_pointless(aes(colour = after_stat(location)), location = "all", size = 3) +
  labs(subtitle = "orientation = 'y'")
```

When `location = "all"`, points are drawn bottom to top in the order:
`"maximum"` < `"minimum"` < `"last"` < `"first"`. Passing an explicit vector
lets you override this:

```{r pointless-order, fig.height = 3, fig.show = 'hold'}
df2 <- data.frame(x = 1:2, y = 1:2)
p2  <- ggplot(df2, aes(x, y)) + geom_path() + coord_equal() +
       scale_colour_manual(values = loc_cols)

p2 + geom_pointless(aes(colour = after_stat(location)),
       location = c("first", "last", "minimum", "maximum"), size = 4) +
     labs(subtitle = "first on top")

p2 + geom_pointless(aes(colour = after_stat(location)),
       location = c("maximum", "minimum", "last", "first"), size = 4) +
     labs(subtitle = "maximum on top")
```

### Facets

`geom_pointless()` respects ggplot2's group structure and works naturally with
facets:

```{r pointless-facets, fig.height = 7}
ggplot(economics_long, aes(x = date, y = value)) +
  geom_line(colour = "grey60") +
  geom_pointless(
    aes(colour = after_stat(location)),
    location = c("minimum", "maximum"),
    size     = 2
  ) +
  stat_pointless(location = "minimum", geom = "text", aes(label = after_stat(y)), vjust = -1) +
  scale_colour_manual(
    values = c(minimum = "#f4ae1b", maximum = "#311dfc")
  ) +
  facet_wrap(vars(variable), ncol = 1, scales = "free_y") +
  theme(legend.position = "bottom") +
  labs(x = NULL, y = NULL, colour = NULL)
```

### Using a different geom with stat_pointless

`stat_pointless()` can drive any geom. Here it adds horizontal reference lines
at the minimum and maximum:

```{r pointless-hline, fig.height = 3}
set.seed(42)
df3 <- data.frame(x = 1:10, y = sample(10))

ggplot(df3, aes(x, y)) +
  geom_line() +
  stat_pointless(
    aes(yintercept = y, colour = after_stat(location)),
    location = c("minimum", "maximum"),
    geom     = "hline"
  ) +
  scale_colour_manual(values = c(minimum = "#f4ae1b", maximum = "#311dfc")) +
  theme(legend.position = "bottom")
```
