---
title: "Introduction to ggpointless"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ggpointless}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  dev       = "ragg_png",
  fig.align = "center",
  out.width = "100%"
)
library(ggplot2)
library(ggpointless)

roma <- scico::scico(4, palette = "roma")
theme_set(
  theme_minimal() +
    theme(
      palette.colour.discrete = \(n) scico::scico(n, palette = "roma"),
      palette.fill.discrete   = \(n) scico::scico(n, palette = "roma"),
      palette.fill.continuous = scales::colour_ramp(scico::scico(256, palette = "roma")),
      geom                    = element_geom(fill = "#7E1700")
    )
)
```

`ggpointless` is a small extension of the [`ggplot2`](https://ggplot2.tidyverse.org/)
package that provides additional layers:

1. `geom_area_fade()`
2. `geom_arch()` & `stat_arch()`
3. `geom_catenary()` & `stat_catenary()`
4. `geom_chaikin()` & `stat_chaikin()`
5. `geom_fourier()` & `stat_fourier()`
6. `geom_lexis()` & `stat_lexis()`
7. `geom_point_glow()`
8. `geom_pointless()` & `stat_pointless()`


## geom_area_fade

`geom_area_fade()` behaves like `geom_area()` but fills each group with a
linear gradient that fades from the fill colour to transparent at the baseline,
drawing the eye to the signal rather than the area underneath.

```{r area-fade-economics, fig.height = 3}
ggplot(economics, aes(date, unemploy)) +
  geom_area_fade()
```

Use `alpha` to set the starting opacity and `alpha_fade_to` to control where
the gradient ends. The direction can be reversed — transparent at the top,
opaque at the baseline — by swapping their values.

```{r area-fade-alpha, fig.height = 3}
ggplot(economics, aes(date, unemploy)) +
  geom_area_fade(alpha = 0.5, alpha_fade_to = 0.05)
```

Multiple groups each get their own gradient:

```{r area-fade-groups, fig.height = 3}
df <- data.frame(
  g = c("a", "a", "a", "b", "b", "b"),
  x = c(1, 3, 5, 2, 4, 6),
  y = c(2, 5, 1, 3, 6, 7)
)
ggplot(df, aes(x, y, fill = g)) +
  geom_area_fade(alpha = 0.5)
```


### 2D gradient

When `fill` is mapped to a variable inside `aes()`, ggplot2 builds a
**horizontal** colour gradient across each group. `geom_area_fade()` overlays
its vertical alpha schedule on top, giving a true two-dimensional gradient:
colour varies left-to-right while opacity fades from the data line down to
the baseline.

```{r area-fade-2d, fig.height = 3}
set.seed(42)
df_fade <- data.frame(
  x = seq_len(60),
  y = cumsum(rnorm(60, sd = 0.35))
)
df_fade$y <- df_fade$y - mean(df_fade$y)

ggplot(df_fade, aes(x, y, fill = y)) +
  geom_area_fade(colour = "#333333")
```

### Device compatibility and the fallback

The 2D gradient depends on **Porter-Duff compositing**, a feature of R's
graphics engine added in R 4.2. When the active graphics device does not
support compositing (e.g. `grDevices::pdf()`), `geom_area_fade()` falls
back to a single-colour vertical fade: the horizontal colour gradient is
lost, only the vertical alpha fade survives, and a one-time message is
emitted.

Most modern raster devices — including `ragg::agg_png()` and the
Cairo-backed `png()` shipped on Linux and macOS — **do** support
compositing, so the 2D gradient works out of the box. In RStudio, go to
*Tools > Global Options > Graphics > Backend* and select *AGG* to ensure
full support.

## geom_arch & stat_arch

### Catenary curves and arches

A **catenary** is the curve formed by a flexible chain or cable hanging freely
between two fixed supports. It follows this equation:

$$y = a \cosh\!\left(\frac{x}{a}\right)$$

`geom_arch()` draws these inverted catenary curves between successive points.
The shape is controlled by `arch_length` or `arch_height` (vertical rise above
the highest endpoint of each segment). By default the arch length is twice
the Euclidean distance.

```{r arch-basic, fig.height = 3}
ggplot(data.frame(x = 1:2, y = c(0, 0)), aes(x, y)) +
  geom_arch(arch_height = 0.6)
```

### The Rice House

A real-world application is the **[Rice House](https://en.wikipedia.org/wiki/Rice_House,_Eltham)**
in Eltham, whose distinctive roofline consists of shallow catenary arched
spans rising above vertical columns. The sketch below reconstructs its simplified
cross-section:

```{r rice-house, fig.height = 3}
rice_house <- data.frame(
  x = c(0, 1.5, 2.5, 3.5, 5),
  y = c(0, 1, 1, 1, 0)
)

ggplot(rice_house, aes(x, y)) +
  geom_arch(arch_height = 0.15, linewidth = 2, colour = "#333333") +
  geom_segment(aes(xend = x, yend = 0), colour = "#333333") +
  geom_hline(yintercept = 0, colour = "#4a7c59", linewidth = 3) +
  coord_equal() +
  theme_void() +
  labs(caption = "Rice House, Eltham (simplified catenary cross-section)")
```

### Controlling arch height with stat_arch

`stat_arch()` exposes the underlying computation. Here it shows the effect of
different `arch_height` values between the same two endpoints:

```{r stat-arch-compare, fig.height = 3.5}
ggplot(data.frame(x = c(0, 2), y = c(0, 0)), aes(x, y)) +
  stat_arch(arch_height = 0.5, colour = roma[1]) +
  stat_arch(arch_height = 1.5, colour = roma[2]) +
  stat_arch(arch_height = 3.0, colour = roma[3]) +
  labs(title = "arch_height = 0.5, 1.5, 3.0")
```


## geom_catenary & stat_catenary

`geom_catenary()` draws the *non*-inverted, hanging-chain form between
successive points.

```{r catenary-basic, fig.height = 3}
set.seed(1)
ggplot(data.frame(x = 1:6, y = sample(6)), aes(x, y)) +
  geom_catenary() +
  geom_point(size = 3, colour = "#333333")
```

### chain_length

By default the chain length is twice the Euclidean distance per segment. Pass
`chain_length` to set an explicit arc length — longer values make the chain sag
more. If the value is shorter than the straight-line distance, a warning is
issued and a straight line is drawn instead.

```{r catenary-chain-length, fig.height = 3}
ggplot(data.frame(x = c(0, 2), y = c(1, 1)), aes(x, y)) +
  lapply(seq(1.1, 2.5, by = 0.3), \(cl) {
    geom_catenary(chain_length = cl)
  }) +
  ylim(NA, 1.05) +
  labs(title = "Increasing chain_length adds more sag")
```

### sag

`sag` sets the vertical drop below the lowest endpoint of each segment, giving
you direct control over how much the chain droops:

```{r catenary-sag, fig.height = 3}
df_sag <- data.frame(x = c(0, 2, 4, 6), y = c(1, 1, 1, 1))

ggplot(df_sag, aes(x, y)) +
  geom_catenary(sag = c(0.1, 0.5, 1.2)) +
  geom_point(size = 3, colour = "#333333") +
  labs(title = "sag = 0.1, 0.5, 1.2 (left to right)") +
  ylim(c(-.25, 1))
```

### Combining chain_length and sag

Both `chain_length` and `sag` can be specified at the same time. Where both are
given for a segment, `sag` takes precedence. Use `NA` to fall back to the other
argument for that segment:

```{r catenary-sag-vs-chain-length, fig.height = 3}
df_sag <- data.frame(x = c(0, 2, 4, 6), y = c(1, 1, 1, 1))

ggplot(df_sag, aes(x, y)) +
  geom_catenary(
    chain_length = c(2, NA, 3),
    sag = c(0.1, 0.5, NA)
  ) +
  geom_point(size = 3, colour = "#333333") +
  labs(title = "sag wins over chain_length") +
  ylim(c(-.25, 1))
```

## geom_chaikin & stat_chaikin

Chaikin's corner-cutting algorithm smooths a polygonal path by iteratively
replacing each corner with two new points placed at a given ratio along the
adjacent edges. With `ratio = 0.25` (the default) and `iterations = 5`, the
path converges to a smooth B-spline approximation.

```{r chaikin-basic, fig.height = 3}
set.seed(42)
dat <- data.frame(x = seq.int(10), y = sample(15:30, 10))

ggplot(dat, aes(x, y)) +
  geom_line(linetype = "dashed", colour = "#333333") +
  geom_chaikin(colour = roma[2])
```

### Effect of ratio

`ratio` controls how aggressively corners are cut. At `ratio = 0.5`, the new
points coincide with the edge midpoints, producing the maximum rounding
possible in a single pass.

```{r chaikin-ratio, fig.height = 3}
triangle <- data.frame(x = c(0, 0.5, 1), y = c(0, 1, 0))

ggplot(triangle, aes(x, y)) +
  geom_polygon(fill = NA, colour = "grey70", linetype = "dashed") +
  geom_chaikin(ratio = 0.10, mode = "closed", colour = roma[1]) +
  geom_chaikin(ratio = 0.25, mode = "closed", colour = roma[2]) +
  geom_chaikin(ratio = 0.50, mode = "closed", colour = roma[3]) +
  coord_equal() +
  labs(subtitle = "ratio = 0.10, 0.25, 0.50")
```

### Smoothing a closed polygon with stat_chaikin

`stat_chaikin()` can be combined with any geom. Pass `geom = "polygon"` to
smooth the boundary of a filled shape:

```{r chaikin-polygon, fig.height = 3.5}
# An irregular hexagon
hex <- data.frame(
  x = c(0.0, 0.4, 1.0, 1.2, 0.9, 0.1),
  y = c(0.2, 1.0, 0.9, 0.3, -0.2, 0.0)
)

ggplot(hex, aes(x, y)) +
  geom_polygon(fill = "grey85", colour = "#333333", linetype = "dashed") +
  stat_chaikin(geom = "polygon", mode = "closed", alpha = 0.45, colour = NA) +
  coord_equal() +
  labs(title = "Original polygon (dashed) with smoothed fill (purple)")
```


## geom_fourier & stat_fourier

`geom_fourier()` fits a truncated **discrete Fourier transform** to the
supplied x/y data and renders the reconstructed curve. Fewer harmonics produce
a smoother, low-frequency summary; retaining all harmonics reproduces the
original signal exactly (up to interpolation artefacts).

```{r fourier-harmonics, fig.height = 3.5}
set.seed(42)
n <- 150
df_f <- data.frame(
  x = seq(0, 2 * pi, length.out = n),
  y = sin(seq(0, 2 * pi, length.out = n)) +
    0.4 * sin(3 * seq(0, 2 * pi, length.out = n)) +
    rnorm(n, sd = 0.25)
)

ggplot(df_f, aes(x, y)) +
  geom_point(alpha = 0.25, size = 0.8, colour = "#333333") +
  geom_fourier(n_harmonics = 1, colour = roma[1]) +
  geom_fourier(n_harmonics = 3, colour = roma[2]) +
  geom_fourier(n_harmonics = 10, colour = roma[3]) +
  labs(
    title = "1 harmonic, 3, 10 (light to dark)",
    x = NULL, y = NULL
  )
```

### Detrending

A linear drift in the data will dominate the low-frequency coefficients,
preventing the Fourier series from capturing the periodic structure. Set
`detrend = "lm"` (or `"loess"`) to subtract the trend before the transform;
it is added back afterwards so the output remains on the original scale.

```{r fourier-detrend, fig.height = 3.5}
set.seed(3)
x_d <- seq(0, 4 * pi, length.out = 100)
df_d <- data.frame(
  x = x_d,
  y = sin(x_d) + x_d * 0.4 + rnorm(100, sd = 0.2)
)

ggplot(df_d, aes(x, y)) +
  geom_point(alpha = 0.35, size = 0.8, colour = "#333333") +
  geom_fourier(aes(colour = "as is"),
    n_harmonics = 3,
    linetype = "dashed"
  ) +
  geom_fourier(aes(colour = "detrend = 'lm'"), n_harmonics = 3, detrend = "lm") +
  labs(
    title = "geom_fourier() with/without detrending",
    x = NULL, y = NULL
  )
```

### Pitfall: irregular spacing with calendar dates

The Fourier transform (via `stats::fft()`) assumes that observations are
**evenly spaced** in time. When you supply calendar dates directly, this
assumption is silently violated: months have 28–31 days, so consecutive
observations are not the same distance apart.

The example below maps `date` — a `Date` vector — to the x aesthetic and
asks `geom_fourier()` to reconstruct the signal with all harmonics
(`n_harmonics = NULL`). You might expect the Fourier curve to pass through
every data point, but it does not:

```{r fourier-dates-fail, fig.height = 3.5}
ggplot(head(economics, 12 * 6), aes(date, psavert)) +
  geom_line(linetype = "dotted") +
  geom_point(size = 1) +
  geom_fourier(n_harmonics = NULL) +
  labs(
    title = "geom_fourier() with Date x-axis — curve misses observations",
    x = NULL
  )
```

The Fourier layer converts dates to their underlying integer representation
(days since 1970-01-01), constructs a **uniform** grid of that many integer
steps, computes the inverse FFT on that grid, and draws a line through those
interpolated points. Because the grid points fall between observation dates,
the curve misses the actual data points even when every harmonic is included.

To fix this you can replace the date x-axis with a plain integer
index so that spacing is guaranteed to be uniform. You can restore
human-readable year labels with `scale_x_continuous()`:

```{r fourier-dates-fix, fig.height = 3.5}
eco <- head(economics, 12 * 6)
eco$tmp <- seq_len(nrow(eco))

# integer positions for year-start labels
yr_breaks <- which(format(eco$date, "%m") == "01")

ggplot(eco, aes(tmp, psavert)) +
  geom_line(linetype = "dotted") +
  geom_point(size = 1) +
  geom_fourier(n_harmonics = NULL) +
  scale_x_continuous(
    breaks = yr_breaks,
    labels = format(eco$date[yr_breaks], "%Y")
  )
```

The Fourier curve now interpolates the data exactly. When working with any
irregularly-spaced time series — monthly, weekly, or business-day — convert
the time variable to a sequential integer index before passing it to
`geom_fourier()` or `stat_fourier()`.

## geom_lexis & stat_lexis

`geom_lexis()` draws a 45° lifeline for each observation from its start to its
end. The required aesthetics are `x` and `xend`; `y` and `yend` are calculated
by `stat_lexis()` and represent the cumulative duration.

```{r lexis-basic, fig.height = 3.5}
df_l <- data.frame(
  key  = c("A", "B", "B", "C", "D"),
  x    = c(0, 1, 6, 5, 6),
  xend = c(5, 4, 10, 8, 10)
)

p <- ggplot(df_l, aes(x = x, xend = xend, colour = key)) +
  coord_equal()

p + geom_lexis()
```

When there is a gap between two events of the same cohort, a horizontal dotted
segment bridges the gap. Set `gap_filler = FALSE` to hide it.

```{r lexis-gap, fig.height = 3.5}
p + geom_lexis(gap_filler = FALSE)
```

### Using after_stat(type) and custom point styling

The computed variable `type` takes the value `"solid"` for 45° lifelines and
`"dotted"` for horizontal gap-fillers. Map it to `linetype` to make the
distinction explicit, and use `point_colour` to style the endpoint dot
independently of the lifeline colour.

```{r lexis-type, fig.height = 3.5}
p +
  stat_lexis(
    aes(linetype = after_stat(type)),
    point_colour = "#333333",
    shape        = 21,
    fill         = "white",
    size         = 2.5,
    stroke       = 0.8
  ) +
  scale_linetype_identity()
```

### Date and POSIXct classes

`geom_lexis()` works with Date and POSIXct objects as well as numerics. The
y-axis shows duration in the native unit of the scale (days for Date, seconds
for POSIXct).

```{r lexis-dates, fig.height = 3.5}
df_dates <- data.frame(
  key   = c("A", "B"),
  start = c(2019, 2021),
  end   = c(2022, 2022)
)
df_dates[, c("start", "end")] <- lapply(
  df_dates[, c("start", "end")],
  function(i) as.Date(paste0(i, "-01-01"))
)

ggplot(df_dates, aes(x = start, xend = end, group = key)) +
  geom_lexis() +
  scale_y_continuous(
    breaks = 0:3 * 365.25,
    labels = \(i) paste0(floor(i / 365.25), " yr")
  ) +
  coord_fixed() +
  labs(y = "Duration")
```


## geom_point_glow

`geom_point_glow()` is a drop-in replacement for `geom_point()` that adds a
radial gradient glow behind each point using `grid::radialGradient()`. It
shines brightest on dark backgrounds.

### Ursa Major

The constellation Ursa Major, drawn with star positions in right ascension and
declination:

```{r ursa-major, fig.height = 5, fig.width = 7.5}
# Precise J2000 coordinates (Hipparcos / SIMBAD) and apparent magnitudes
# from the Wikipedia Großer Bär star table.
stars_base <- data.frame(
  star = c(
    "Dubhe", "Merak", "Phecda", "Megrez", "Alioth", "Mizar", "Alkaid",
    "Muscida", "23 UMa", "Upsilon", "Phi", "Theta",
    "Talitha", "Alkaphrah",
    "Chi", "Psi", "Tania Borealis", "Tania Australis",
    "Alula Borealis", "Alula Australis"
  ),
  ra_h = c(
    11.062, 11.031, 11.897, 12.257, 12.900, 13.399, 13.792,
    8.504, 9.526, 9.850, 9.868, 9.548,
    8.987, 9.060,
    11.767, 11.161, 10.285, 10.372,
    11.308, 11.303
  ),
  dec_d = c(
    61.75, 56.38, 53.70, 57.03, 55.96, 54.93, 49.31,
    60.72, 63.06, 59.04, 54.06, 51.68,
    48.04, 47.16,
    47.78, 44.50, 42.91, 41.50,
    33.09, 31.53
  ),
  # Apparent visual magnitudes (Wikipedia / Bright Star Catalogue).
  # Alioth is slightly variable; 1.77 is the mean.
  mag = c(
    1.81, 2.34, 2.41, 3.32, 1.77, 2.23, 1.86,
    3.35, 3.65, 3.78, 4.55, 3.20,
    3.12, 3.57,
    3.69, 3.00, 3.45, 3.06,
    3.49, 3.79
  )
)

# Linear size mapping: brighter (lower mag) → larger point
mag_to_size <- function(m) pmax(0.7, (5.5 - m) * 1.0)

# Constellation figure lines (IAU official figure):
# Big Dipper (bowl closed, then handle)
path1 <- c("Alkaid", "Mizar", "Alioth", "Megrez", "Phecda", "Merak", "Dubhe", "Megrez")
# Head + neck → front shoulder junction
path2 <- c("Dubhe", "23 UMa", "Muscida", "Upsilon", "Theta")
# Belly (Merak) to front-leg junction
path3 <- c("Merak", "Upsilon")
# Front paw pair
path4 <- c("23 UMa", "Upsilon", "Theta", "Alkaphrah", "Talitha")
# Hind leg → Alula paw pair
path5 <- c("Phecda", "Chi", "Alula Borealis", "Alula Australis")
# Hind leg branch → Tania paw pair
path6 <- c("Chi", "Psi", "Tania Australis", "Tania Borealis")

make_path <- function(star_names, group_id) {
  df <- stars_base[match(star_names, stars_base$star), ]
  df$group <- group_id
  df
}

plot_data <- rbind(
  make_path(path1, 1),
  make_path(path2, 2),
  make_path(path3, 3),
  make_path(path4, 4),
  make_path(path5, 5),
  make_path(path6, 6)
)
plot_data$x <- -plot_data$ra_h
plot_data$y <- plot_data$dec_d

ggplot(plot_data, aes(x = x, y = y)) +
  geom_path(aes(group = group), colour = "#F5F5F5", linewidth = 0.6, alpha = .6) +
  geom_point_glow(
    data        = stars_base,
    aes(x = -ra_h, y = dec_d, size = mag_to_size(mag)),
    shape       = 8,
    colour      = ggplot2::alpha("#F5F5F5", alpha = .6),
    glow_colour = "#F5F5F5",
    glow_alpha  = 0.75
  ) +
  scale_size_identity() +
  geom_text(
    data = stars_base,
    aes(x = -ra_h, y = dec_d, label = star),
    colour = "#bbccdd",
    vjust = -1.2,
    size = 2.5,
    check_overlap = TRUE
  ) +
  scale_x_continuous(
    breaks = seq(-14, -8, by = 1),
    labels = \(x) paste0(abs(x), "h")
  ) +
  scale_y_continuous(labels = \(x) paste0(x, "°")) +
  labs(
    title = "Ursa Major",
    x     = NULL,
    y     = NULL
  ) +
  theme(
    panel.background = element_rect(fill = grid::linearGradient(colours = c("#081849", "#070836")), colour = NA),
    plot.background  = element_rect(fill = grid::linearGradient(colours = c("#081849", "#070836")), colour = NA),
    panel.grid.major = element_line(colour = "#1e2f3e", linetype = "dotted"),
    panel.grid.minor = element_blank(),
    text             = element_text(colour = "#344B73"),
    plot.title       = element_text(size = 14, face = "bold")
  )
```


## geom_pointless & stat_pointless

`geom_pointless()` highlights selected observations — first, last, minimum,
maximum, or all four — by default with points. Its name reflects that it is not
particularly useful on its own, but in conjunction with `geom_line()` and
friends, it adds useful context at a glance.

```{r pointless-basic, fig.height = 3}
x <- seq(-pi, pi, length.out = 100)
df1 <- data.frame(var1 = x, var2 = rowSums(outer(x, 1:5, \(x, y) sin(x * y))))

p <- ggplot(df1, aes(x = var1, y = var2)) +
  geom_line()

p + geom_pointless(location = "all", size = 3)
```

Map the computed variable `location` to `colour` to distinguish the four roles:

```{r pointless-colour, fig.height = 3}
p +
  geom_pointless(
    aes(colour = after_stat(location)),
    location = "all",
    size = 3
  ) +
  theme(legend.position = "bottom")
```

### Order and orientation

Locations are determined in data order, as `geom_path()` would — not sorted by
x like `geom_line()`. This matters when the path crosses itself:

```{r pointless-spiral, fig.height = 3.5, fig.show = 'hold'}
x <- seq(5, -1, length.out = 1000) * pi
spiral <- data.frame(var1 = sin(x) * 1:1000, var2 = cos(x) * 1:1000)

p_spi <- ggplot(spiral) +
  geom_path() +
  coord_equal(xlim = c(-1000, 1000), ylim = c(-1000, 1000)) +
  theme(legend.position = "none")

p_spi +
  aes(x = var1, y = var2) +
  geom_pointless(aes(colour = after_stat(location)), location = "all", size = 3) +
  labs(subtitle = "orientation = 'x'")

p_spi +
  aes(y = var1, x = var2) +
  geom_pointless(aes(colour = after_stat(location)), location = "all", size = 3) +
  labs(subtitle = "orientation = 'y'")
```

When `location = "all"`, points are drawn bottom to top in the order:
`"maximum"` < `"minimum"` < `"last"` < `"first"`. Passing an explicit vector
lets you override this:

```{r pointless-order, fig.height = 3, fig.show = 'hold'}
df2 <- data.frame(x = 1:2, y = 1:2)
p2 <- ggplot(df2, aes(x, y)) +
  geom_path() +
  coord_equal()

p2 + geom_pointless(aes(colour = after_stat(location)),
  location = c("first", "last", "minimum", "maximum"), size = 4
) +
  labs(subtitle = "first on top")

p2 + geom_pointless(aes(colour = after_stat(location)),
  location = c("maximum", "minimum", "last", "first"), size = 4
) +
  labs(subtitle = "maximum on top")
```

`geom_pointless()` respects `ggplot2`'s group structure and works naturally with
facets:

```{r pointless-facets, fig.height = 7}
ggplot(
  subset(economics_long, variable %in% c("psavert", "unemploy")),
  aes(x = date, y = value)
) +
  geom_line(colour = "#333333") +
  geom_pointless(
    aes(colour = after_stat(location)),
    location = c("minimum", "maximum"),
    size = 2
  ) +
  stat_pointless(location = "minimum", geom = "text", aes(label = after_stat(y)), hjust = -1) +
  facet_wrap(vars(variable), ncol = 1, scales = "free_y") +
  theme(legend.position = "bottom") +
  labs(x = NULL, y = NULL, colour = NULL)
```

Here it adds horizontal reference lines at the minimum and maximum:

```{r pointless-hline, fig.height = 3}
set.seed(42)
df3 <- data.frame(x = 1:10, y = sample(10))

ggplot(df3, aes(x, y)) +
  geom_line() +
  stat_pointless(
    aes(yintercept = y, colour = after_stat(location)),
    location = c("minimum", "maximum"),
    geom = "hline"
  ) +
  theme(legend.position = "bottom")
```
