---
title: "Introduction to ggpointless"
author: "Markus Döring"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ggpointless}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # out.width = "100%",
  dev = "ragg_png" # <3
)
library(ggpointless)
```


## Add some points

```{r lib}
library(ggplot2)
library(ggpointless)
```





```{r default, fig.show='hold'}
library(ggpointless)
theme_set(theme_minimal())

cols <- c('#f4ae1b', '#d77e7b', '#a84dbd', '#311dfc')

x <- seq(-pi, pi, length.out = 100)
y <- outer(x, 1:5, FUN = \(x, y) sin(x*y)) |> 
  rowSums()

df1 <- data.frame(
  var1 = x,
  var2 = y
)

p <- ggplot(df1, aes(x = var1, y = var2)) + coord_equal()
p + geom_pointless() + labs(subtitle = "default")
p + geom_pointless(location = "all") + labs(subtitle = "location = 'all'")
```

As you can see, `geom_pointless()` is not terribly useful on its own (and here I stuck thinking about a better package name) but it is when it teams up with `geom_line()`, hopefully. `geom_pointless()` understands the same arguments as `geom_point()` and one additional argument `location`. You can set it to `"first"`, `"last"` (default), `"minimum"`, `"maximum"`, and `"all"`, where `"all"` is just shorthand to select `"first"`, `"last"`, `"minimum"` and `"maximum"`.  

The functions compute one additional variable -- `location` -- that you can map to an aesthetic, e.g. `color`.  

```{r location, fig.show='hold'}
p <- p + geom_line()
p + geom_pointless(location = "all",
                        size = 3)

p + geom_pointless(aes(colour = after_stat(location)),
                   location = "all",
                   size = 3) +
  scale_color_manual(values = cols)
```

### Order and orientation

The order of your data is determined by how the observations appear in the data -- like `geom_path()` does compared to `geom_line()`. This can be seen in the next example, the example data is taken from the [`geomtextpath`](https://github.com/AllanCameron/geomtextpath/#using-geomtextpath) package:

```{r spiral, fig.show='hold'}
x <- seq(5, -1, length.out = 1000) * pi
spiral <- data.frame(var1 = sin(x) * 1:1000, 
                     var2 = cos(x) * 1:1000)

p <- ggplot(spiral) +
  coord_equal() +
  scale_color_manual(values = cols) +
  theme(legend.position = "none")


p + aes(x = var1, y = var2) +
  geom_path() + 
  geom_pointless(aes(colour = after_stat(location)),
                 location = "all") +
  labs(subtitle = "orientation = 'x'")

p + aes(y = var1, x = var2) +
  geom_path() +
  geom_pointless(aes(colour = after_stat(location)),
                   location = "all") +
  labs(subtitle = "orientation = 'y'")
```

As you see from the last examples on the left hand side `"first"` and `"minimum"` overlap, and `"first"` wins over `"minimum"`. The order in which points are plotted (if specified together) from top to bottom is: `"first"` > `"last"` > `"minimum"` > `"maximum"`. 


## Data

The `ggpointless` package contains two dataset:

1. `co2_ml` : [CO~2~ records taken at Mauna Loa](https://gml.noaa.gov/ccgg/trends/data.html)
2. `covid_vac` : [COVID-19 Cases and Deaths by Vaccination Status](https://covid.cdc.gov/covid-data-tracker/#rates-by-vaccine-status)

