---
title: "Introduction to ggpointless"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ggpointless}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "ragg_png" # <3
)
library(ggpointless)
```

`ggpointless` is a small extension of the [`ggplot2`](https://ggplot2.tidyverse.org/) that provides two additional layers:

1. `geom_pointless()` / `stat_pointless()`
2. `geom_lexis()`

and 

## geom_pointless 
`geom_pointless()` is a layer to easily add minimal emphasis to your plots. The function takes it's power from `stat_pointless()`, which actually does all the work, but is not usually in the spotlight.

```{r default_pointless}
library(ggplot2)
library(ggpointless)

x <- seq(-pi, pi, length.out = 100)
y <- outer(x, 1:5, function(x, y) sin(x * y))

df1 <- data.frame(
  var1 = x,
  var2 = rowSums(y)
)

p <- ggplot(df1, aes(x = var1, y = var2))
p + geom_pointless(location = c("first", "last", "minimum", "maximum"))
```

As you see, just adding `geom_pointless()` to `ggplot(...)` is not terribly useful on its own but when it teams up with `geom_line()` and friends, hopefully.


```{r location}
p <- p + geom_line()
p + geom_pointless(location = "all", size = 3)
```


`geom_pointless()` behaves like `geom_point()` does with the addition of a `location` argument. You can set it to `"first"`, `"last"`, the default, `"minimum"`, `"maximum"`, and `"all"`; where `"all"` is just shorthand to select `"first"`, `"last"`, `"minimum"` and `"maximum"`.  

In addition, you can use the computed variable `location` and map it to an aesthetic, e.g. `color`.  

```{r color}
p + geom_pointless(aes(color = after_stat(location)),
  location = "all",
  size = 3
) +
  theme(legend.position = "bottom")
```

### Order and orientation

The locations are determined in the order in which they appear in the data, like `geom_path()` does compared to `geom_line()`. This can be seen in the next example, with sample data kindly taken from the [`geomtextpath`](https://github.com/AllanCameron/geomtextpath/#using-geomtextpath) package:

```{r spiral, fig.show='hold'}
x <- seq(5, -1, length.out = 1000) * pi
spiral <- data.frame(
  var1 = sin(x) * 1:1000,
  var2 = cos(x) * 1:1000
)

p <- ggplot(spiral) +
  geom_path() +
  coord_equal(xlim = c(-1000, 1000), ylim = c(-1000, 1000)) +
  theme(legend.position = "none")

p + aes(x = var1, y = var2) +
  geom_pointless(aes(color = after_stat(location)),
    location = "all",
    size = 3
  ) +
  labs(subtitle = "orientation = 'x'")

p + aes(y = var1, x = var2) +
  geom_pointless(aes(color = after_stat(location)),
    location = "all",
    size = 3
  ) +
  labs(subtitle = "orientation = 'y'")
```

As you see from the first of the last two examples `"first"` and `"minimum"` overlap, and `"first"` wins over `"minimum"`. If `location` is set to `"all"`, then the order in which points are plotted from top to bottom is: `"first"` > `"last"` > `"minimum"` > `"maximum"`. 

Otherwise, the order is determined as specified in the `location` argument, which also applies to the order of the legend key labels.

```{r overplotting, fig.show='hold'}
df2 <- data.frame(
  var1 = 1:2,
  var2 = 1:2
)
p <- ggplot(df2, aes(x = var1, y = var2)) +
  geom_path() +
  coord_equal()

# same as location = 'all'
p + geom_pointless(aes(color = after_stat(location)),
  location = c("first", "last", "minimum", "maximum"),
  size = 3
) +
  labs(subtitle = "same as location = 'all'")
# reversed order
p + geom_pointless(aes(color = after_stat(location)),
  location = c("maximum", "minimum", "last", "first"),
  size = 3
) +
  labs(subtitle = "custom order")
# same as location = 'all' again
p + geom_pointless(aes(color = after_stat(location)),
  location = c("maximum", "minimum", "last", "first", "all"),
  size = 3
) +
  labs(subtitle = "same as location = 'all' again")
```


### Pick another geom

Just like all `stat_*` functions, `stat_pointless()` has a default geom, which is `"point"`. This means in reverse that you can highlight e.g. minimum and maximum in another way, for example with a horizontal line.

```{r stat}
set.seed(42)
ggplot(data.frame(x = 1:10, y = sample(1:10)), aes(x, y)) +
  geom_line() +
  stat_pointless(
    aes(yintercept = y, color = after_stat(location)),
    location = c("minimum", "maximum"),
    geom = "hline"
  ) +
  guides(color = guide_legend(reverse = TRUE))
```

## geom_lexis

The idea of the `geom_lexis()` function is to draw a lifeline for an event from it's start to it's end. The required aesthetics are `x` and `xend`. 

```{r default_lexis}
df1 <- data.frame(
  key = c("A", "B", "B", "C", "D"),
  start = c(0, 1, 6, 5, 6),
  end = c(5, 4, 10, 8, 10)
)

p <- ggplot(df1, aes(x = start, xend = end, color = key)) +
  coord_equal()
p + geom_lexis()
```

If there is a gap in an event, this is shown by a horizontal line, that you can hide. Also, `geom_lexis()` creates an additional variable, `type`, you can map to e.g. the linetype aesthetic (or any other aesthetic) using `after_scale()`.

```{r after_scale, fig.show='hold'}
p + geom_lexis(gap_filler = FALSE)
p + geom_lexis(aes(linetype = after_scale(type)))
```


## Data

The `ggpointless` package contains two data sets:

1. `co2_ml` : [CO~2~ records taken at Mauna Loa](https://gml.noaa.gov/ccgg/trends/data.html)
2. `covid_vac` : [COVID-19 Cases and Deaths by Vaccination Status](https://covid.cdc.gov/covid-data-tracker/#rates-by-vaccine-status)

See the [`vignette("examples")`]() for possible use cases.
