---
title: "Doppelgänger"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Doppelgänger}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  dev       = "ragg_png",
  dpi       = 300,
  fig.asp   = 0.618,
  fig.width = 6,
  fig.height = 3,
  fig.align = "center",
  out.width = "80%"
)
```

```{=html}
<style>
  .content h3 {
    margin-top: -30px !important;
  }

  details {
    margin-bottom: 40px;
  }
</style>
```

All `geom_*_fade()` functions in this vignette are drop-in replacements for
their ggplot2 counterparts. So is `geom_point_glow()`. They expect the same data,
support the same aesthetics, and use the same defaults — with two
small additions:

1. linear or radial gradient effect, and for rectangular geoms:
2. support for rounded corners

They do not add additional statistical meaning.

All effects use either `grid::linearGradient()`, `grid::radialGradient()` or
[Porter-Duff compositing](https://www.stat.auckland.ac.nz/~paul/Reports/GraphicsEngine/compositing/compositing.html). 

> **Note**: Most but not all graphic devices that ship with R support 
> gradients (ragg, Cairo-backed png on Linux/macOS, svg). The notable
> exception is `grDevices::pdf()`.

```{r setup-theme, warning=FALSE}
library(ggpointless)

# set consistent theme for all plots
cols <- c("#311dfc", "#a84dbd", "#d77e7b", "#f4ae1b")
theme_set(
  theme_minimal() +
    theme(legend.position = "bottom") +
    theme(geom = element_geom(fill = cols[1])) +
    theme(palette.fill.discrete = c(cols[1], cols[3])) +
    theme(palette.colour.discrete = cols)
  )
```

## geom_area_fade

`geom_area_fade()` behaves like `geom_area()` but fills each group with a
linear gradient that fades from the fill colour to transparent at the baseline at
`y = 0`.

```{r area-fade-economics}
ggplot(economics, aes(date, unemploy)) +
  geom_area_fade()
```

Use `alpha` to set the starting opacity and `alpha_fade_to` to control at which
alpha value the gradient ends.

```{r area-fade-alpha, fig.height = 3}
ggplot(economics, aes(date, unemploy)) +
  geom_area_fade(alpha = 0.75, alpha_fade_to = 0.1)
```

The direction can be reversed — transparent at the top, opaque at the
baseline — by swapping their values. The outline colour is unaffected from
the alpha logic.

```{r geom-area-fade-reverse, warning=FALSE}
ggplot(economics, aes(date, unemploy)) +
  geom_area_fade(alpha = 0, alpha_fade_to = 1)
```

### 2D gradient

When `fill` is mapped to a variable inside `aes()`, ggplot2 builds a
**horizontal** colour gradient across each group. `geom_area_fade()` overlays
its vertical alpha schedule on top, giving a true two-dimensional gradient:
colour varies left-to-right while opacity fades from the data line down to
the baseline.

```{r area-fade-2d, fig.height = 3}
set.seed(42)
ggplot(economics, aes(date, unemploy)) +
  geom_area_fade(aes(fill = uempmed), colour = cols[1]) +
  scale_fill_continuous(palette = scales::colour_ramp(cols))
```

### Device compatibility and the fallback

The 2D gradient depends on **Porter-Duff compositing**^[See: https://www.stat.auckland.ac.nz/~paul/Reports/GraphicsEngine/compositing/compositing.html], a feature of R's graphics engine added in R 4.2. When the active graphics
device does not support compositing (e.g. `grDevices::pdf()`),
`geom_area_fade()` falls back to a single-colour vertical fade: the
horizontal colour gradient is lost, only the vertical alpha fade survives, and a one-time message is
emitted:

> ! `geom_area_fade()`: the graphics device does not support gradient fills.
> The `fill` colour gradient is replaced by a single colour. Switch to a device that supports gradients (e.g. `ragg::agg_png()`, `svg()`, `cairo_pdf()`) for the full effect.
> This message is displayed once per session.

Most modern raster devices — including `ragg::agg_png()` and the
Cairo-backed `png()` shipped on Linux and macOS — **do** support
compositing, so the 2D gradient works out of the box. In RStudio, go to
*Tools > Global Options > Graphics > Backend* and select *AGG* to ensure
full support.

### Multiple groups

When your data contains multiple groups those are stacked (`position = "stack"`)
and aligned (`stat = "align"`) -- just like `geom_area()` does it. By default,
the alpha fade scales to the global maximum across _all_ groups
(`alpha_scope = "global"`), so equal `|y|` always maps to equal opacity.

```{r geom-area-fade-multiple-groups-basic, warning=FALSE}
df1 <- data.frame(
  g = c("a", "a", "a", "b", "b", "b"),
  x = c(1, 3, 5, 2, 4, 6),
  y = c(2, 5, 1, 3, 6, 7)
)

ggplot(df1, aes(x, y, fill = g)) +
  geom_area_fade()
```

When groups have very different amplitudes or you may not use the default
`position = "stack"` but `stat = "identity"` instead, this can make smaller
groups nearly invisible next to dominant groups.

```{r geom-area-fade-global, warning=FALSE}
df_alpha_scope <- data.frame(
  g = c("a", "a", "a", "b", "b", "b"),
  x = c(1, 3, 5, 2, 4, 6),
  y = c(1, 2, 1, 9, 10, 8)
)
p <- ggplot(df_alpha_scope, aes(x, y, fill = g))
p + geom_area_fade(
  alpha_scope = "global", # default
  position = "identity"
)
```

Setting `alpha_scope = "group"` lets the algorithm calculate the alpha range
for each group separately.

```{r geom-area-fade-group, warning=FALSE}
p <- ggplot(df_alpha_scope, aes(x, y, fill = g))

# alpha_scope = "group": each group uses the alpha range independently
p + geom_area_fade(
  alpha_scope = "group",
  position = "identity"
  )
```

## geom_col_fade

`geom_col_fade()` is the fading counterpart of `geom_col()`: each bar is filled
with a vertical alpha gradient, opaque at the tip and transparent at the
baseline. `geom_bar_fade()` is the `stat = "count"` alias. Both accept a
`radius` argument to draw bars with rounded corners and an `alpha_scope`
argument to control whether the gradient is computed per bar (`"bar"`), per
x-position (`"group"`), or relative to the tallest bar in the panel (`"global"`).

```{r col-fade-dodge, fig.height = 4}
# Mean highway MPG by car class and drive type
df_hwy <- aggregate(hwy ~ class + drv, data = mpg, FUN = mean)
df_hwy <- subset(df_hwy, drv %in% c("f", "r"))

ggplot(df_hwy, aes(reorder(class, hwy), hwy, fill = drv)) +
  geom_col_fade(
    position    = "dodge",
    alpha_scope = "global",
    radius      = grid::unit(5, "pt")
  ) +
  scale_fill_manual(
    values = c(f = cols[1], r = cols[3]),
    labels = c(f = "front-wheel", r = "rear-wheel")
  ) +
  labs(
    x    = NULL,
    fill = NULL
    )
```

## geom_density_fade

`geom_density_fade()` pairs `GeomAreaFade` with `stat_density()` to compute and
draw a kernel density estimate with a fading gradient. All smoothing parameters
(`bw`, `adjust`, `kernel`, `bounds`, …) are forwarded to the stat.

```{r density-fade-iris, fig.height = 3.5}
ggplot(iris, aes(Sepal.Length, fill = Species, colour = Species)) +
  geom_density_fade(position = "identity", alpha = 0.75) +
  labs(x = "Sepal length (cm)", y = NULL, fill = NULL, colour = NULL)
```

### geom_ridgeline_density_fade

This is a wrapper for [geom_ridgeline_fade]() where `stat` parameter
uses `"density"` by default. A ridgeline plot, popularised by the cover of
Joy Division's album [Unknown Pleasures](https://en.wikipedia.org/wiki/Unknown_Pleasures),
stacks multiple lines vertically along the y-dimension. 

Credit to the [ggridges packages](https://wilkelab.org/ggridges/index.html).

```{r density-ridges-fade-iris, fig.height = 3.5}
ggplot(iris, aes(Sepal.Length, y = Species)) +
  geom_ridgeline_density_fade() +
  labs(y = NULL)
```

## geom_freqpoly_fade

`geom_freqpoly_fade()` draws the area under a frequency polygon — the filled
region between the binned line and the baseline — using the same fading gradient
as `geom_area_fade()`. It is paired with `stat_bin()`, so all binning parameters
(`bins`, `binwidth`, `center`, `boundary`, …) are forwarded.

```{r freqpoly-fade-mpg, fig.height = 3.5}
ggplot(mpg, aes(hwy, fill = drv, colour = drv)) +
  geom_freqpoly_fade(position = "identity", alpha = 0.7, bins = 20) +
  scale_fill_manual(
    values = cols[1:3],
    labels = c("4" = "four-wheel", "f" = "front-wheel", "r" = "rear-wheel")
  ) +
  scale_colour_manual(
    values = cols[1:3],
    labels = c("4" = "four-wheel", "f" = "front-wheel", "r" = "rear-wheel")
  ) +
  labs(x = "Highway MPG", y = NULL, fill = NULL, colour = NULL)
```

## geom_histogram_fade

`geom_histogram_fade()` is the fading counterpart of `geom_histogram()`:
rounded, gradient-filled bars via `GeomColFade` paired with `stat_bin()`.
`alpha_scope = "global"` scales all bars relative to the tallest one in the
panel, so the height of a bar is reflected twice — in its physical size and
in how opaque it is.

```{r histogram-fade-diamonds, fig.height = 3.5}
ggplot(diamonds, aes(carat, fill = cut)) +
  geom_histogram_fade(alpha_scope = "global", bins = 40, colour = NA) +
  scale_fill_viridis_d(direction = -1) +
  labs(x = "Carat", y = "Count", fill = NULL)
```

## geom_point_glow

`geom_point_glow()` is a drop-in replacement for `geom_point()` that adds a
radial gradient glow behind each point using `grid::radialGradient()`. By default,
alpha, colour and size inherit their values from `geom_point()`.

```{r point-glow-basic, fig.height = 3.5}
# Basic usage
ggplot(mtcars, aes(wt, mpg, colour = factor(cyl))) +
   geom_point_glow()
```

You can control the alpha, colour, and size of the gradient with these arguments:

- `glow_alpha`
- `glow_colour`
- `glow_size`


```{r point-glow-overwrite-params, fig.height = 3.5}
# Customizing glow parameters (fixed for all points)
ggplot(mtcars, aes(wt, mpg, colour = factor(cyl))) +
  geom_point_glow(
    glow_alpha = 0.25,
    glow_colour = "#0833F5",
    glow_size = 5
    )
```


### Big Dipper

The constellation [Big Dipper](https://en.wikipedia.org/wiki/Big_Dipper), drawn with star positions in right ascension and declination.

```{r big-dipper, fig.height = 5, fig.width = 7.5, echo = FALSE}
# source: https://de.wikipedia.org/wiki/Gro%C3%9Fer_B%C3%A4r#Sterne
# colours, coordinates all approximations of course
big_dipper <- data.frame(
  star = c(
    "Megrez",
    "Dubhe",
    "Merak",
    "Phecda",
    "Megrez",
    "Alioth",
    "Mizar",
    "Alkaid"
  ),
  ra_h = c(12.257, 11.062, 11.031, 11.897, 12.257, 12.900, 13.399, 13.792),
  dec_d = c(57.03, 61.75, 56.38, 53.70, 57.03, 55.96, 54.93, 49.31),
  mag = c(3.32, 1.81, 2.34, 2.41, 3.32, 1.77, 2.23, 1.86),
  colour = c("#CFDDFF", "#FFDBBF", "#C7D9FF", "#C8D9FF", "#CFDDFF", "#C7D9FF", "#CBDBFF", "#BAD0FF")
)

big_dipper$x <- -big_dipper$ra_h
big_dipper$y <- big_dipper$dec_d

# Linear size mapping: brighter (lower mag) → larger point
mag_to_size <- \(m) pmax(0.7, (5.5 - m) * 1.0)

ggplot(big_dipper, aes(x = x, y = y)) +
  geom_path(colour = "#F5F5F5", linewidth = 0.6, alpha = .6) +
  geom_point_glow(
    data = big_dipper[-1L, ], # don't plot Megrez a second time
    aes(x = -ra_h, y = dec_d, size = mag_to_size(mag), colour = colour,
        alpha = mag),
    shape = 8,
    glow_alpha = 0.75
  ) +
  scale_alpha_continuous(range = c(1, 0.4), guide = "none") +
  scale_size_identity() +
  scale_colour_identity() +
  geom_text(
    aes(x = -ra_h, y = dec_d, label = star),
    colour = "#bbccdd",
    vjust = -1.5,
    size = 2.5,
    check_overlap = TRUE
  ) +
  scale_x_continuous(
    breaks = seq(-14, -8, by = 1),
    labels = \(x) paste0(abs(x), "h")
  ) +
  scale_y_continuous(labels = \(x) paste0(x, "°")) +
  labs(
    title = "Big Dipper",
    x = NULL,
    y = NULL
  ) +
  coord_cartesian(clip = 'off') +
  theme(
    panel.background = element_rect(fill = grid::linearGradient(colours = c("#1D2180", "#081849")), colour = NA),
    plot.background = element_rect(fill = grid::linearGradient(colours = c("#1D2180", "#081849")), colour = NA),
    panel.grid = element_blank(),
    text = element_text(colour = "#344B73"),
    plot.title = element_text(size = 18, face = "bold")
  )
```

```{r big-dipper-code, eval = FALSE}
# source: https://de.wikipedia.org/wiki/Gro%C3%9Fer_B%C3%A4r#Sterne
# colours, coordinates all approximations of course
big_dipper <- data.frame(
  star = c(
    "Megrez",
    "Dubhe",
    "Merak",
    "Phecda",
    "Megrez",
    "Alioth",
    "Mizar",
    "Alkaid"
  ),
  ra_h = c(12.257, 11.062, 11.031, 11.897, 12.257, 12.900, 13.399, 13.792),
  dec_d = c(57.03, 61.75, 56.38, 53.70, 57.03, 55.96, 54.93, 49.31),
  mag = c(3.32, 1.81, 2.34, 2.41, 3.32, 1.77, 2.23, 1.86),
  colour = c("#CFDDFF", "#FFDBBF", "#C7D9FF", "#C8D9FF", "#CFDDFF", "#C7D9FF", "#CBDBFF", "#BAD0FF")
)

big_dipper$x <- -big_dipper$ra_h
big_dipper$y <- big_dipper$dec_d

# Linear size mapping: brighter (lower mag) → larger point
mag_to_size <- \(m) pmax(0.7, (5.5 - m) * 1.0)

ggplot(big_dipper, aes(x = x, y = y)) +
  geom_path(colour = "#F5F5F5", linewidth = 0.6, alpha = .6) +
  geom_point_glow(
    data = big_dipper[-1L, ], # don't plot Megrez a second time
    aes(x = -ra_h, y = dec_d, size = mag_to_size(mag), colour = colour,
        alpha = mag),
    shape = 8,
    glow_alpha = 0.75
  ) +
  scale_alpha_continuous(range = c(1, 0.4), guide = "none") +
  scale_size_identity() +
  scale_colour_identity() +
  geom_text(
    aes(x = -ra_h, y = dec_d, label = star),
    colour = "#bbccdd",
    vjust = -1.5,
    size = 2.5,
    check_overlap = TRUE
  ) +
  scale_x_continuous(
    breaks = seq(-14, -8, by = 1),
    labels = \(x) paste0(abs(x), "h")
  ) +
  scale_y_continuous(labels = \(x) paste0(x, "°")) +
  labs(title = "Big Dipper", x = NULL, y = NULL) +
  coord_cartesian(clip = 'off') +
  theme(
    panel.background = element_rect(fill = grid::linearGradient(colours = c("#1D2180", "#081849")), colour = NA),
    plot.background = element_rect(fill = grid::linearGradient(colours = c("#1D2180", "#081849")), colour = NA),
    panel.grid = element_blank(),
    text = element_text(colour = "#344B73"),
    plot.title = element_text(size = 18, face = "bold")
  )
```

## geom_rect_fade

The `geom_rect_fade()` function draws a rectangle with a linear colour gradient,
allowing you to subtly highlight elements. In the following example, a
`geom_segment_fade()` is also added to serve as a border. Like all
`geom_*_fade()` functions, this one is primarily intended to enhance the visual
appeal for the viewer.


```{r rect-fade, fig.height = 3.5}
df <-data.frame(
  x = seq_len(15),
  y = c(90, 93, 88, 91, 86, 85, 89, 84, 96, 72, 68, 66, 62, 59, 46)
)

ggplot(df, aes(x, y)) +
  geom_rect_fade(
    data = data.frame(xmin = 10, xmax = 15, ymin = 0, ymax = 100),
    mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = cols[1],
    alpha = 0,
    alpha_fade_to = 0.1,
    inherit.aes = FALSE
  ) +
  geom_segment_fade(
    data = data.frame(x = 10, xend = 10, y = 100, yend = 0),
    aes(x = x, xend = xend, y = y, yend = yend),
    colour = cols[1],
    alpha = 0.75,
    alpha_fade_to = 0,
    inherit.aes = FALSE
  ) +
  geom_fourier(colour = cols[3]) +
  geom_point(size = 4, colour = cols[3], alpha = 0.25) +
  scale_x_continuous(NULL, breaks = 10, labels = "Did something\nchange here?") +
  scale_y_continuous(NULL, limits = c(0, NA), labels = \(x) paste(x, "%")) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
    theme(
    axis.text.x = element_text(
      size = 10,
      colour = ggplot2::alpha(cols[1], 0.75),
      face = "bold"
    )
  )
```

## geom_ridgeline_fade

`geom_ridgeline_fade()` draws overlapping ridge shapes — each group's area sits
at its own vertical baseline — with the same fading gradient as
`geom_area_fade()`. The convenience wrapper `geom_ridgeline_density_fade()` pairs
the geom with `stat_density()` to compute kernel density estimates on the fly.



### Women in UK Headlines

The chart below is inspired by [*Women in Headlines*](https://pudding.cool/2022/02/women-in-headlines/)
by The Pudding (Jan Diehm and Amber Thomas, 2022). The original piece tracks how
often words describing women appear in news headlines across four countries from
2010 to 2021. The data are fetched directly from The Pudding's public data
repository.

```{r ridgeline-pudding-data, message=FALSE}
url_pudding <- paste0(
  "https://raw.githubusercontent.com/the-pudding/data/master/",
  "women-in-headlines/country_time_freqrank.csv"
)
df_women <- read.csv(url_pudding)

# UK only; drop the "all countries" aggregate rows
df_uk <- subset(df_women, country == "UK")

# Top 14 words by headline-frequency rank (lower rank = more prominent)
top_words <- unique(df_uk[, c("word", "hfreq_rank")])
top_words <- top_words[!duplicated(top_words$word), ]
top_words <- head(top_words[order(top_words$hfreq_rank), ], 14)

df_plot <- subset(df_uk, word %in% top_words$word)

# Order factor: most prominent word at top (highest y)
word_order <- rev(top_words$word[order(top_words$hfreq_rank)])
df_plot$word <- factor(df_plot$word, levels = word_order)
df_plot$y_pos <- as.numeric(df_plot$word)

# Normalise height per word: each ridge's own peak = 1
# so all waves are visually comparable regardless of absolute frequency
maxima <- tapply(df_plot$freq_prop_headlines, df_plot$word, max, na.rm = TRUE)
maxima[maxima == 0] <- 1
df_plot$height_norm <- df_plot$freq_prop_headlines /
  maxima[as.character(df_plot$word)]
```

```{r ridgeline-pudding-plot, fig.height=7, fig.asp=NULL}
ggplot(
  df_plot,
  aes(year, y_pos,
    height = height_norm,
    group  = word,
    fill   = word
  )
) +
  geom_ridgeline_fade(
    scale        = 0.85,
    alpha_fade_to = 0.05,
    alpha_scope  = "area",
    colour       = NA
  ) +
  scale_x_continuous(
    breaks = seq(2010, 2021, by = 2),
    expand = c(0.02, 0)
  ) +
  scale_y_continuous(
    breaks = seq_along(levels(df_plot$word)),
    labels = levels(df_plot$word),
    expand = c(0.02, 0)
  ) +
  scale_fill_viridis_d(option = "C", direction = -1, guide = "none") +
  labs(
    title    = "Women in UK Headlines, 2010\u20132021",
    subtitle = paste0(
      "Each wave shows how often a word appeared relative to its own peak year.\n",
      "Words ranked by headline frequency (top = most prominent)."
    ),
    caption  = paste0(
      "Data: The Pudding \u2014 Jan Diehm & Amber Thomas (2022), ",
      "pudding.cool/2022/02/women-in-headlines/\n",
      "Reproduced with geom_ridgeline_fade() from ggpointless."
    ),
    x = NULL,
    y = NULL
  ) +
  theme(
    plot.background  = element_rect(fill = "#0d1117", colour = NA),
    panel.background = element_rect(fill = "#0d1117", colour = NA),
    panel.grid.major.x = element_line(colour = "#ffffff22", linewidth = 0.3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    text               = element_text(colour = "#e0e0e0"),
    axis.text          = element_text(colour = "#aaaaaa", size = 8),
    plot.title         = element_text(colour = "#ffffff", face = "bold", size = 13),
    plot.subtitle      = element_text(colour = "#aaaaaa", size = 8, lineheight = 1.4),
    plot.caption       = element_text(colour = "#555555", size = 6.5, lineheight = 1.3),
    plot.margin        = margin(12, 16, 10, 12)
  )
```

## geom_unit_col

`geom_unit_col()` renders each value as a vertical strip of discrete cells —
one cell per unit — coloured by `fill`. It follows the same `x`/`y` convention
as `geom_col()`: `x` identifies each bar; `y` is a pre-computed integer count per
segment. Use `geom_unit_bar()` when you want ggplot2 to count observations for
you (like `geom_bar()`).

### Title Towns

The chart below is inspired by [*Title Towns*](https://pudding.cool/2018/11/titletowns/)
by The Pudding (Russell Goldenberg, 2018), which visualises how many professional
sports championships each US city has won. Each cell represents one title.
With `coord_flip()`, each cell occupies one unit along the count axis; the flat-strip
appearance comes from having many cells per bar, not from non-square cells.

```{r col-waffle-titletowns, fig.asp = 0.55, fig.width = 7}
titles <- data.frame(
  city = rep(
    c("New York", "Boston", "Chicago", "Los Angeles",
      "Detroit", "San Francisco", "Philadelphia", "Pittsburgh"),
    each = 4
  ),
  sport = rep(c("NFL", "NBA", "NHL", "MLB"), times = 8),
  n = c(
     9,  2, 11, 35,   # New York
     5, 17,  6, 10,   # Boston
    10,  6,  6,  6,   # Chicago
     2, 11,  3,  6,   # Los Angeles
     4,  3, 11,  4,   # Detroit
     7,  4,  0,  7,   # San Francisco
     5,  4,  2,  7,   # Philadelphia
     6,  0,  5,  5    # Pittsburgh
  )
)

# Order cities top-to-bottom by total title count
totals       <- tapply(titles$n, titles$city, sum)
titles$city  <- factor(titles$city, levels = names(sort(totals)))
titles$sport <- factor(titles$sport, levels = c("MLB", "NBA", "NFL", "NHL"))
titles       <- titles[titles$n > 0L, ]

cols_sport <- c(
  "MLB" = "#30D840",
  "NBA" = "#3EA3DB",
  "NFL" = "#2E46BC",
  "NHL" = "#ECE833"
)

ggplot(titles, aes(x = city, y = n, fill = sport)) +
  geom_unit_col() +
  scale_fill_manual(values = cols_sport) +
  coord_flip() +
  labs(
    title    = "Title Towns",
    subtitle = "Major professional sports championships by US city (NFL, NBA, NHL, MLB)",
    caption  = paste0(
      "Inspired by \u2018Title Towns\u2019 \u00b7 The Pudding (Russell Goldenberg, 2018)\n",
      "pudding.cool/2018/11/titletowns/ \u00b7 ",
      "Reproduced with geom_unit_col() from ggpointless"
    ),
    x    = NULL,
    fill = NULL
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.text.y        = element_text(size = 9),
    plot.caption       = element_text(size = 6.5, lineheight = 1.3),
    legend.position    = "bottom"
  )
```
